---
title: "HW 5"
author: "Jacob Haussler"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
`chemical.(2)` <- read.csv("./chemical.csv")
#Dat = read.csv("C:/Users/Brad/Desktop/Stat 4020/chemical.csv")
Dat <- `chemical.(2)`
```


### 1. Fit a multiple regression model relating CO2 product ( y ) to total solvent ( x 6 ) and hydrogen consumption ( x 7 ). Write the equation of the fitted regression line.
```{r}
mlr <- lm(y ~ x6+x7, data = Dat)
summary(mlr)
```
- The equation of the fitted regression line is: Y(hat) = 2.526460 + 0.018522x6 + 2.185753x7

### 2. Perform an influential analysis and clearly state your findings.

```{r}
k=2
n=27
hv <- hatvalues(mlr)
which(hv > (2*(k+1)/n))
row_num <- c(1:27)
plot(row_num, hv, xlab = "row number",
     ylab = "hatvalues", 
     main = "Identification of high leverage points")
abline(h = (2*(k+1)/n))
```

- Here we can see that the 19th is a high leverage point in the data.

### Question 2 continued
```{r}
r1 <- rstandard(mlr)
which(abs(r1) > 3)
row_num <- c(1:27)
plot(c(1,27), c(-6, 6), xlab = "row number",
     ylab = "internally studentized residuals", 
     main = "Identification of outlier points",
     type = "n")
points(row_num, r1)
abline(h = 3)
abline(h = -3)
```
- This shows no case of an outlier because residuals greater than 3 are considered outleirs and none of these residuals are greater than 3. 

### Question 2 continued
```{r}
cd <- cooks.distance(mlr) 
which(cd > 0.5)
row_num <- c(1:27)
plot(row_num, cd, xlab = "row number",
     ylab = "Cook's distance", 
     main = "Identification of influential points")
abline(h = 0.5)
```
- Typically Di > 0.5 indicates that the data point could be influential. Since the 17th and 26th data points are above 0.5. We can say that they could be influential. 

We're going to check for influential points using other methods also.

COVRATIO
```{r}
cv = covratio(mlr)
plot(cv, main="Influential Points")

h0 = 1 + 3*(k+1)/n
h1 = 1 - 3*(k+1)/n
which(cv > h0)
which(cv < h1)
abline(h=h0)
abline(h=h1)
```

using COVRATIO shows that points 10, 17, 19, 26, and 27 may be influential also.

DFFITS:
```{r}
dffit = dffits(mlr)
h2 = 2*sqrt((k+2)/(n-k-2))
which(dffit > h2)
plot(dffit, main="Influential Points")
abline(h=h2)
```

DFFITS shows only point 26 as an influential point in the graph. 

Based on the above methods, 26 is considered an influential point in all of them. Other points that may need to be considered as influential are:
10, 17, 19, 27.

